### 2013-07-05 Tex/chapters/results.tex
Weitere Ausfuehrung zur autokorrelation, todo: plot tau uber T

For evaluation the Monte Carlo simulation is run until the system
is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
and the magnetisation per site \(m=\frac{1}{N}\sum_i s_i\) and energy
per site \(E=\frac{1}{N}\hat H\) are calculated and saved for every
\(2\tau\) sweeps. Where \(\tau\) denotes the \emph{autocorrelation time}
(see also \ref{ssec:eqtime}).
For every observable \(O\) the expected value \(\avg{O}\) is determined
as the mean of \(10000\) measurements for \(L=16,32\) or \(5000\) for
\(L=64,128\). The expected values \(\avg{O}\) for \(100\) different
random proximity graphs with the same disorder paramter \(\sigma\)
are then averaged to \(\overline{\avg{O}}\). The errors are estimated
by bootstrapping \cite{Bootstrap}.

@@

        \subfigure[][]
        {
            \label{sfig:Tc_norm:RNG}
            \includegraphics[width=0.45\textwidth]{plots/RNG_Tc_norm}
        }
        \subfigure[][]
        {
            \label{sfig:Tc_norm:GG}
            \includegraphics[width=0.45\textwidth]{plots/GG_Tc_norm}
        }
        \caption[Critical Temperature over different disorder parameters]
        {
            Top: Critical temperatures over different
            disorder parameters for
            \subref{sfig:Tc:RNG} the Relative Neighborhood Graph and
            \subref{sfig:Tc:GG} the Gabriel Graph.\\
            Middle: Mean sum of the coupling constants to all
            neighbors over different disorder parameters for
            \subref{sfig:sumJ:RNG} the Relative Neighborhood Graph and
            \subref{sfig:sumJ:GG} the Gabriel Graph.\\
            Bottom: Normalized critical temperatures over different
            disorder parameters for
            \subref{sfig:Tc_norm:RNG} the Relative Neighborhood Graph and
            \subref{sfig:Tc_norm:GG} the Gabriel Graph.
        }
        \label{fig:Tc}
    \end{figure}\\
    % I don't know what all that means...
    The Gabriel graph, which is shown on the right side of fig. \ref{fig:Tc},
    jumps from \(\sigma = 0\) to \(\sigma > 0\). This is easily explained
    by the definition of the Gabriel graph and it's aftereffects for
    the transition from \(\sigma = 0\). Visible in fig. \ref{fig:GG_sigma}
    a small change of sigma causes many new egdes to arise\footnote{See also \url{http://www.youtube.com/watch?v=PcVZ2pG11GI} for an animation.}.
    To fully understand this, take four nodes forming a square. The edge
    across does not exist, because the other two nodes are on the edge
    of the lune. Moving one node slightly into the square, causes the lune
    to get smaller, hence no other nodes are inside or on the edge of
    the lune anymore and the edge appears.
    \begin{figure}[htbp]
        \centering
        \subfigure[][]
        {
            \label{sfig:GG_sigma:zero}
            \includegraphics[width=0.40\textwidth]{images/GG/sigma_e0}
        }
        \subfigure[][]
        {
            \label{sfig:GG_sigma:notzero}
            \includegraphics[width=0.40\textwidth]{images/GG/sigma_g0}
        }
        \caption[]
        {
            Gabriel graph with periodic boundary conditions for
                \subref{sfig:GG_sigma:zero} \(\sigma = 0\)
                \subref{sfig:GG_sigma:notzero} \(\sigma = 0.01\)
        }
        \label{fig:GG_sigma}
    \end{figure}\\
    Moreover fig. \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}\subref{sfig:Tc_norm:GG}
    shows monotonically decreasing \(T_c\) with increasing disorder
    parameter \(\sigma\). The forms of both curves are similar, but the
    one for the Relative Neighborhood graph \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}
    is generally lower and spans over a bigger temperature range than
    the curve of the Gabriel graph \ref{fig:Tc}\subref{sfig:Tc_norm:GG}.
    % Warum??? Die eigenschaften des Graphen sollten mit der Normalisierung erledigt sein.
    % RNG: laengere Kanten -> weniger lokale effekte?
    For big \(\sigma\) the curve approaches the limit of randomly
    distributed nodes, hence \(T_c\) is independet of \(\sigma\) for
    \(\sigma >> 1\).\\
    Also both graph types have a plateau at \(0 < \sigma < 0.1\). So
    small disorder has little influence on the normalized critical temperature.
    But once the degree of the graph and hence the mean sum of the coupling
    constants decreases, the normalized critical temperature decreases, too.
    % The author is not sure what it means.
    % mit theo ergebnissen von Honeycomb \cite{Wannier1945} vergleichen (fur deg = 3)
    % cosh(2L_c)=2 mit L_c = J/2kT_c -> T_c \approx 1.52

\subsection{Critical Exponents}
    For \(\sigma \in \{0,0.1,0.2,0.3,0.5,1.0\}\) a finite size scaling analysis was
    performed to determine the critical exponents \(\beta, \gamma, \nu\)
    using \texttt{autoscale.py} \cite{autoscale2009}. The values for
    \(\sigma = 0\) are analytically known \cite{Pelissetto2002}. The
    values for all other \(\sigma\) should be the same as for \(\sigma = 0\)
    like mentioned before in section \ref{ssec:finitesize}.
    The values for \(\sigma\) are chosen to represent every interesting
    region from fig. \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}\subref{sfig:Tc_norm:GG}:
    the plateau at small \(\sigma\), the steep decline and the plateau at
    big \(\sigma\).\\
    Like in tab. \ref{tab:critExp} to see, most values
    are matching the expectations. Most \(\beta\) seem to be a bit too
    big, but they are close enough to the expectations to be explained
    by the fact that small systems (\(L=32,64\)) were used for the
    analysis.
    Anyway, two critical exponents are sufficient to determine the universality
    class. Therefore this disordered Ising modell is in the same universality
    class as the standard Ising ferromagnet \cite[p. 145]{Katzgraber2011}.
    \begin{table}[htbp]
        \center
        \begin{tabular}{l l l l l}
            \toprule
             & \multicolumn{1}{c}{\(\sigma\)} & \multicolumn{1}{c}{\(\nu\)} & \multicolumn{1}{c}{\(\gamma\)} & \multicolumn{1}{c}{\(\beta\)}\\
            \midrule
            exact (\cite[p. 59]{Pelissetto2002}) & \multicolumn{1}{c}{\(0\)} & \multicolumn{1}{c}{\(1\)} & \multicolumn{1}{c}{\(-\frac{7}{4}\)} & \multicolumn{1}{c}{\(\frac{1}{8}\)}\\
            \midrule
            Gabriel      & 0.0 & 1.008(4) & -1.735(2) & 0.1262(4)\\
                         & 0.1 & 1.02(1)  & -1.744(5) & 0.133(6) \\
                         & 0.3 & 1.000(5) & -1.724(16)& 0.129(12)\\
                         & 0.5 & 1.009(8) & -1.750(12)& 0.125(13)\\
                         & 1.0 & 1.015(22)& -1.743(17)& 0.123(16)\\
            \midrule
            Relative N.  & 0.0 & 1.007(2) & -1.739(2) & 0.130(1) \\
                         & 0.1 & 0.99(1)  & -1.746(5) & 0.133(4) \\
                         & 0.2 & 1.022(17)& -1.756(14)& 0.123(10)\\
                         & 0.5 & 1.002(17)& -1.750(16)& 0.143(13)\\
                         & 1.0 & 1.011(20)& -1.758(16)& 0.138(13)\\
            \bottomrule
        \end{tabular}
        \caption{Critical exponents for different \(\sigma\)}
        \label{tab:critExp}
    \end{table}

>>>>

For evaluation the Monte Carlo simulation is run until the system
is equilibrated after \(t_{eq}\) sweeps. Then the simulation continues
and the magnetisation per site \(m=\frac{1}{N}\sum_i s_i\) and energy
per site \(E=\frac{1}{N}\hat H\) are calculated and saved for every
\(2\tau\) sweeps. Where \(\tau\) denotes the \emph{autocorrelation time}
(see also \ref{ssec:eqtime}). The used values for different system sizes
are listed in tab. \ref{tab:tauAndTeq} in numbers of sweeps. Note that
the \(t_{eq}\) values are generously rounded up to be on the safe side
and the \(\tau\) values are determined as the maximum \(\tau\) over all
simulated temperatures.
\[\tau = \max_T \tau_T\]
\begin{table}[htbp]
    \center
    \begin{tabular}{l r r r r r}
        \toprule
        \(L\)    & 16 &  32 &  64 & 128 & 256\\
        \midrule
        \(\tau\) &  3 &   5 &   7 &  12 &  16\\
        \(t_eq\) & 40 & 100 & 100 & 200 & 600\\
        \bottomrule
    \end{tabular}
    \caption{Autocorrelation Times $\tau$ and Equilibration Times $t_{eq}$}
    \label{tab:tauAndTeq}
\end{table}\\
Obiviously the autocorrelation time \(\tau\) increases with the system
size. In fact it obeys a power law, which is the expected bahaivior as
mentioned in section \ref{ssec:eqtime}. In fig. \ref{fig:autocorr}\subref{fig:autocorr:powerLaw}
the dynamical critical exponent \(z\) is determined as \(z=0.57(4)\).
This exponent is much higher than the expectation. For the Wolff cluster
algorithm at criticality, one would expect \(z=0.25\)\cite{NewmanBarkema1999}.
This suggests that \emph{critical slowing down} is not an issue in this
case and the autocorrelation time is limited by the slow decorrelation
at low temperatures. A direct proof is given by fig. \ref{fig:autocorr}\subref{fig:autocorr:temperatures}
Apparently even parallel tempering can not compensate that. Probably
because only one configuration swap between neighboring temperatures is
performed each sweep. Those neighboring configurations at low temperatures
have a magnetisation per site of \(m \approx \pm 1\), so there is a good
chance, that the exchange does not change \(m\) (in case of same sign)
and thus does not lower the autocorrelation function.
\begin{figure}[htbp]
    \centering
    \subfigure[in dependence on the system size $L$][]
    {
        \label{fig:autocorr:powerLaw}
        \includegraphics[width=0.45\textwidth]{plots/autocorrL}
    }
    \subfigure[in dependence on the temperature $T$][]
    {
        \label{fig:autocorr:temperatures}
        \includegraphics[width=0.45\textwidth]{plots/autocorrT}
    }
    \caption[The autocorrelation time $\tau$]
    {
        Dependece of the autocorrelation time $\tau$ on\\
        \subref{fig:autocorr:powerLaw} the system size $L$\\
        \subref{fig:autocorr:temperatures} the temperature $T$
    }
    \label{fig:autocorr}
\end{figure}\\
For every observable \(O\) the expected value \(\avg{O}\) is determined
as the mean of \(N_{\mathrm{measure}}=10000\) measurements for \(L=16,32\)
or \(N_{\mathrm{measure}}=5000\) for \(L=64,128\). The number of
calculated sweeps totals to
\[N_{\mathrm{sweeps}}=t_{eq}+2\tau N_{\mathrm{measure}}\]
The expected values \(\avg{O}\) for \(100\) different random proximity
graphs with the same disorder paramter \(\sigma\) are then averaged to
\(\overline{\avg{O}}\).
The errors are estimated by bootstrapping \cite{Bootstrap}.

@@

        \subfigure[][]
        {
            \label{sfig:Tc_norm:RNG}
            \includegraphics[width=0.45\textwidth]{plots/RNG_Tc_norm}
        }
        \subfigure[][]
        {
            \label{sfig:Tc_norm:GG}
            \includegraphics[width=0.45\textwidth]{plots/GG_Tc_norm}
        }
        \caption[Critical Temperature over different disorder parameters]
        {
            Top: Critical temperatures over different
            disorder parameters for
            \subref{sfig:Tc:RNG} the Relative Neighborhood Graph and
            \subref{sfig:Tc:GG} the Gabriel Graph.\\
            Middle: Mean sum of the coupling constants to all
            neighbors over different disorder parameters for
            \subref{sfig:sumJ:RNG} the Relative Neighborhood Graph and
            \subref{sfig:sumJ:GG} the Gabriel Graph.\\
            Bottom: Normalized critical temperatures over different
            disorder parameters for
            \subref{sfig:Tc_norm:RNG} the Relative Neighborhood Graph and
            \subref{sfig:Tc_norm:GG} the Gabriel Graph.
        }
        \label{fig:Tc}
    \end{figure}\\
    % I don't know what all that means...
    The Gabriel graph, which is shown on the right side of fig. \ref{fig:Tc},
    jumps from \(\sigma = 0\) to \(\sigma > 0\). This is easily explained
    by the definition of the Gabriel graph and it's aftereffects for
    the transition from \(\sigma = 0\). Visible in fig. \ref{fig:GG_sigma}
    a small change of sigma causes many new egdes to arise\footnote{See also \url{http://www.youtube.com/watch?v=PcVZ2pG11GI} for an animation.}.
    To fully understand this, take four nodes forming a square. The edge
    across does not exist, because the other two nodes are on the edge
    of the lune. Moving one node slightly into the square, causes the lune
    to get smaller, hence no other nodes are inside or on the edge of
    the lune anymore and the edge appears.
    \begin{figure}[htbp]
        \centering
        \subfigure[][]
        {
            \label{sfig:GG_sigma:zero}
            \includegraphics[width=0.40\textwidth]{images/GG/sigma_e0}
        }
        \subfigure[][]
        {
            \label{sfig:GG_sigma:notzero}
            \includegraphics[width=0.40\textwidth]{images/GG/sigma_g0}
        }
        \caption[Anomaly of the Gabriel graph for small $\sigma$]
        {
            Gabriel graph with periodic boundary conditions for
                \subref{sfig:GG_sigma:zero} \(\sigma = 0\)
                \subref{sfig:GG_sigma:notzero} \(\sigma = 0.01\)
        }
        \label{fig:GG_sigma}
    \end{figure}\\
    Moreover fig. \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}\subref{sfig:Tc_norm:GG}
    shows monotonically decreasing \(T_c\) with increasing disorder
    parameter \(\sigma\). The forms of both curves are similar, but the
    one for the Relative Neighborhood graph \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}
    is generally lower and spans over a bigger temperature range than
    the curve of the Gabriel graph \ref{fig:Tc}\subref{sfig:Tc_norm:GG}.
    % Warum??? Die eigenschaften des Graphen sollten mit der Normalisierung erledigt sein.
    % RNG: laengere Kanten -> weniger lokale effekte?
    For big \(\sigma\) the curve approaches the limit of randomly
    distributed nodes, hence \(T_c\) is independet of \(\sigma\) for
    \(\sigma >> 1\).\\
    Also both graph types have a plateau at \(0 < \sigma < 0.1\). So
    small disorder has little influence on the normalized critical temperature.
    But once the degree of the graph and hence the mean sum of the coupling
    constants decreases, the normalized critical temperature decreases, too.
    % The author is not sure what it means.
    % mit theo ergebnissen von Honeycomb \cite{Wannier1945} vergleichen (fur deg = 3)
    % cosh(2L_c)=2 mit L_c = J/2kT_c -> T_c \approx 1.52

\subsection{Critical Exponents}
    For \(\sigma \in \{0,0.1,0.2,0.3,0.5,1.0\}\) a finite size scaling analysis was
    performed to determine the critical exponents \(\beta, \gamma, \nu\)
    using \texttt{autoscale.py} \cite{autoscale2009}. The values for
    \(\sigma = 0\) are analytically known \cite{Pelissetto2002}. The
    values for all other \(\sigma\) should be the same as for \(\sigma = 0\)
    like mentioned before in section \ref{ssec:finitesize}.
    The values for \(\sigma\) are chosen to represent every interesting
    region from fig. \ref{fig:Tc}\subref{sfig:Tc_norm:RNG}\subref{sfig:Tc_norm:GG}:
    the plateau at small \(\sigma\), the steep decline and the plateau at
    big \(\sigma\).\\
    Like in tab. \ref{tab:critExp} to see, most values
    are matching the expectations. Most \(\beta\) seem to be a bit too
    big, but they are close enough to the expectations to be explained
    by the fact that small systems (\(L=32,64\)) were used for the
    analysis.
    Anyway, two critical exponents are sufficient to determine the universality
    class. Therefore this disordered Ising modell is in the same universality
    class as the standard Ising ferromagnet \cite[p. 145]{Katzgraber2011}.
    \begin{table}[htbp]
        \center
        \begin{tabular}{l l l l l}
            \toprule
             & \multicolumn{1}{c}{\(\sigma\)} & \multicolumn{1}{c}{\(\nu\)} & \multicolumn{1}{c}{\(\gamma\)} & \multicolumn{1}{c}{\(\beta\)}\\
            \midrule
            exact (\cite[p. 59]{Pelissetto2002}) & \multicolumn{1}{c}{\(0\)} & \multicolumn{1}{c}{\(1\)} & \multicolumn{1}{c}{\(-\frac{7}{4}\)} & \multicolumn{1}{c}{\(\frac{1}{8}\)}\\
            \midrule
            Gabriel      & 0.0 & 1.008(4) & -1.735(2) & 0.1262(4)\\
                         & 0.1 & 1.02(1)  & -1.744(5) & 0.133(6) \\
                         & 0.3 & 1.000(5) & -1.724(16)& 0.129(12)\\
                         & 0.5 & 1.009(8) & -1.750(12)& 0.125(13)\\
                         & 1.0 & 1.015(22)& -1.743(17)& 0.123(16)\\
            \midrule
            Relative N.  & 0.0 & 1.007(2) & -1.739(2) & 0.130(1) \\
                         & 0.1 & 0.99(1)  & -1.746(5) & 0.133(4) \\
                         & 0.2 & 1.022(17)& -1.756(14)& 0.123(10)\\
                         & 0.5 & 1.002(17)& -1.750(16)& 0.143(13)\\
                         & 1.0 & 1.011(20)& -1.758(16)& 0.138(13)\\
            \bottomrule
        \end{tabular}
        \caption{Critical exponents for different $\sigma$}
        \label{tab:critExp}
    \end{table}
